---
title: "Week 1 Homework"
author: "Andrew Joyce"
date: "01-15-2023"
format: html
editor: visual
mainfont: "Baskerville"
embed-resources: true
toc: true
bibliography: references.bib
---

```{r}
#| echo: false
#| results: hide
#| include: false

library(infer)
library(janitor) 
library(latex2exp)

```

## 1.1 Exercise

| Term                       | Definition                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               |
|-----------------------|-------------------------------------------------|
| DAG (Direct Acyclic Graph) | A visual representation of a causal mechanism, that is, the pathway from cause to effect of various variables. it is "direct" (directly shows the cause and effect order) and "acyclic" (there is a certain path, variables are not allowed to loop or cycle around)                                                                                                                                                                                                                                     |
| Estimand                   | The object of inquiry- the precise quantity about which we marshal data to draw an inference.                                                                                                                                                                                                                                                                                                                                                                                                            |
| Theoretical estimand       | The goal stated *outside* the model. Has two components: a unit-specific quantity and the target population. These combined to create the theroetical estimand: what we would like to know if we had data for the full population in all factual or counterfactual worlds of interest.                                                                                                                                                                                                                   |
| Empirical estimand         | Only observable quantities that we have in the data.                                                                                                                                                                                                                                                                                                                                                                                                                                                     |
| Unit-specific quantity     | Within a theoretical estimand, consists of either a realized outcome (If $i$ is employed), potential outcome (if $i$ is employed after job training), difference in potential outcomes (effect of job training on employment of $i$), or potential outcome that would be realized under intervention of more than one variable (if $i$ is employed if they have job training and childcare). also clarifies if research goal is causal, and if so, what counterfactual intervention is being considered. |

## 1.2 Exercise

```{r}
library(tidyverse)
library(gt)
library(pwr)
```

## 1.3 Exercise

```{r}
d <- data.frame(
  T = c(0, 0, 1, 0, 0, 1, 1, 1),
  Y0 = c(5, 8, 5, 12, 4, 8, 4, 9),
  Y1 = c(5, 10, 3, 13, 2, 9, 1, 13), 
  id = LETTERS[1:8]
)

gt(d, rowname_col = "id") # gt is used for fancy printing of tables
```

> Create a new variable called `Y` that contains the observed outcomes.

```{r}
d <- d |> 
  mutate(Y = if_else(T == 0, Y0, Y1))
```

> What is the Average Treatment Effect (ATE) for this 8 person experiment?

```{r}

mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])

```

The average treatment effect is -0.75 (a reduction of 0.75 units for the treatment group).

## 1.4 Exercise

> -   Simulate a new completely randomized experiment on these 8 people; that is, re sample $T$ at random so that equal numbers get the treatment and the control.
>
> ```{=html}
> <!-- -->
> ```
> -   Create a new variable called `Y` that contains the observed outcomes.
>
> -   What is the Average Treatment Effect (ATE) for this 8 person experiment?

```{r}
d <- d |> 
  mutate(T = sample(d$T)) |> 
  mutate(Y = if_else(T == 0, Y0, Y1))

mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])

d <- d |> 
  mutate(T = sample(d$T)) |> 
  mutate(Y = if_else(T == 0, Y0, Y1))

mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])

d <- d |> 
  mutate(T = sample(d$T)) |> 
  mutate(Y = if_else(T == 0, Y0, Y1))

mean(d$Y[d$T == 1]) - mean(d$Y[d$T == 0])

```

> Do this a couple of times (at least 3) and note the differences.
>
> How do these estimates compare to the “real” ATE?

These $\text{ATE}$ values vary greatly between one another, and differ greatly from the "real" ATE of $0.125$. Some of these simulated values are negative and some are positive. Some are fairly close (in terms of absolute value) to the real ATE, some are much further apart.

## 1.5 Exercise

> Obviously, an experiment of 8 people will not give you enough “statistical power.”
>
> -   Assuming the ATE is $0.125$, how many people would you need to enroll in this experiment to have enough statistical power?

Ultimately, I am doing a two-sample t-test by comparing the mean for the treatment group with the mean for the control group. "Power analysis" can help me determine what sample size I need based on what kind of test I am running. The package `power.t.test` can calculate the needed $N$.

-   $N$: null, since that is the value I am interested in calculating

-   $\delta$: the average treatment effect, 0.125

-   $\text{sig. level}$: the probability of a type I error. Sociologists use $0.05$.

-   $\text{power}$: 1 - probability of type II error. Typically, power = 0.8 since expected probably of a type II error is 20%.

```{r}
power.t.test(n = NULL, delta = 0.125, sd = 1, sig.level = 0.05, power = 0.8, type="two.sample", alternative = "two.sided")

```

We'd need 2012 (1006 \* 2) total people, minus the 8 already enrolled, so we would need a total of 2004 additional people to enroll in the experiment. That's a lot of people to enroll!

> *\*I found a different function, pwr.t.test, that does a similar thing as above. However, I can't specify delta. I can specify d ( here, I'm using the vague "large" effect size with a cohen's D of 0.8).*

```{r}
pwr.t.test(n= NULL, d = 0.8, sig.level=0.05, power = 0.8, type = "two.sample", alternative = "two.sided")
```

But this gives me an n of 26 per group, so a total n of 52, in which we would only need to enroll 44 more subjects. We should discuss the differences between these two approaches- and if this is even the right approach to begin with.
