---
title: "Week 5 Homework"
author: "Andrew Joyce"
date: "02-12-2023"
format: html
editor: visual
mainfont: "Baskerville"
embed-resources: true
toc: true
---

# Setup

```{r}
library(tidyverse)
library(gssr)
library(marginaleffects)
library(broom)

gss2022 <- gss_get_yr(2022)
```

# 5.1 Instructions & mtcars example

```{r}
ols <- lm(mpg ~ disp + am, data = mtcars)

new_am0 <- mtcars |> 
  mutate(am = 0)

new_am1 <- mtcars |> 
  mutate(am = 1)

p0 <- predict(ols, newdata = new_am0) ## predictions for am == 0
p1 <- predict(ols, newdata = new_am1) ## predictions for am == 1

## interestingly
mean(p1) - mean(p0) 
## looks like the value for beta 'am' in ols
```

# 5.2 Linear Regression

```{r}
d <- gss2022 |> 
  select(tvhours, degree, madeg, padeg) |> 
  mutate(pardeg = pmax(madeg, padeg, na.rm = TRUE),
         college = if_else(degree >= 3, 1L, 0L),
         parcol = if_else(pardeg >= 3, 1L, 0L)) |>
  select(tvhours, college, parcol) |> 
  drop_na()
```

## 5.2.1 Exercise

```{r}
mod1 <- lm(tvhours ~ college + parcol, data = d)

# ATE estimate
avg_slopes(mod1, variables = "college") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(mod1, newdata = new_col0) ## predictions for college == 0
p1 <- predict(mod1, newdata = new_col1) ## predictions for college == 1

# checking results here
mean(p0)
mean(p1)

## finally, ATE estimates
mean(p1) - mean(p0)
mean(p1 - p0) #oh, this does the same thing. makes sense
```

```{r}
# ATT/ATU estimate
avg_slopes(
  model = mod1, 
  variables = "college",
  by = "college" # separately by treatment group
) |> 
  tidy()
```

### ***Manual ATT estimate***

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(mod1, newdata = new_col0)[d$college == 1] ## predictions for college == 0 for treated
p1 <- predict(mod1, newdata = new_col1)[d$college == 1] ## predictions for college == 1 for treated

# ATT
mean(p1 - p0)
```

### ***Manual ATU estimate***

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(mod1, newdata = new_col0)[d$college == 0] ## predictions for college == 0 for untreated
p1 <- predict(mod1, newdata = new_col1)[d$college == 0] ## predictions for college == 1 for untreated

# ATU
mean(p1 - p0)
```

## 5.2.2 Exercise

```{r}
mod2 <- lm(tvhours ~ college * parcol, data = d)

# ATE estimate
avg_slopes(mod2, variables = "college") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(mod2, newdata = new_col0) ## predictions for college == 0
p1 <- predict(mod2, newdata = new_col1) ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

```{r}
# ATT/ATU estimate
avg_slopes(
  model = mod2, 
  variables = "college",
  by = "college" # separately by treatment group
) |> 
  tidy()
```

### *Manual ATT estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(mod2, newdata = new_col0)[d$college == 1] ## predictions for college == 0 for treated
p1 <- predict(mod2, newdata = new_col1)[d$college == 1] ## predictions for college == 1 for treated

# ATT
mean(p1 - p0)
```

### *Manual ATU estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(mod2, newdata = new_col0)[d$college == 0] ## predictions for college == 0 for untreated
p1 <- predict(mod2, newdata = new_col1)[d$college == 0] ## predictions for college == 1 for untreated

# ATU
mean(p1 - p0)
```

# 5.3 Poisson Regression

```{r}
d <- gss2022 |>
  filter(wrkstat == 1) |> # full time workers
  select(realrinc, degree, madeg, padeg, sex, age) |> 
  mutate(pardeg = pmax(madeg, padeg, na.rm = TRUE),
         college = if_else(degree >= 3, 1L, 0L),
         parcol = if_else(pardeg >= 3, 1L, 0L),
         female = if_else(sex == 2, 1L, 0L),
         realrinc = floor(realrinc)) |>             # integer
  select(realrinc, college, parcol, female, age) |> 
  drop_na()
```

## 5.3.1 Exercise

### *Manual ATE Estimate*

```{r}
# marginal effects ATE
qp1 <- glm(realrinc ~ college + (parcol + female + age + I(age^2)), 
           data = d,
           family = "quasipoisson")

avg_slopes(qp1,
           variables = "college",
           type = "link") |> 
  tidy()
```

```{r}
# Manual ATE
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp1, newdata = new_col0) ## predictions for college == 0
p1 <- predict(qp1, newdata = new_col1) ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT Estimate*

```{r}
# marginal effects
avg_slopes(qp1,
           variables = "college",
           type = "link",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
# manual ATT
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp1, newdata = new_col0)[d$college == 1] ## predictions for college == 0 for treated
p1 <- predict(qp1, newdata = new_col1)[d$college == 1] ## predictions for college == 1 for treated

# ATT
mean(p1 - p0)
```

### *Manual ATU estimate*

```{r}
# manual ATU
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp1, newdata = new_col0)[d$college == 0] ## predictions for college == 0 for untreated
p1 <- predict(qp1, newdata = new_col1)[d$college == 0] ## predictions for college == 1 for untreated

# ATU
mean(p1 - p0)
```

## 5.3.2 Exercise

```{r}
avg_slopes(qp1,
           variables = "college",
           type = "response") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp1, newdata = new_col0, type = "response") ## predictions for college == 0
p1 <- predict(qp1, newdata = new_col1, type = "response") ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT estimate*

```{r}
# marginal effects ATT/ATU
avg_slopes(qp1,
           variables = "college",
           type = "response",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp1, newdata = new_col0, type = "response")[d$college == 1] 
p1 <- predict(qp1, newdata = new_col1, type = "response")[d$college == 1] 

# ATT estimate
mean(p1 - p0) 
```

### Manual ATU estimate

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp1, newdata = new_col0, type = "response")[d$college == 0] 
p1 <- predict(qp1, newdata = new_col1, type = "response")[d$college == 0] 

# ATU estimate
mean(p1 - p0) 
```

## 5.3.3 Exercise

### *Manual ATE Estimate*

```{r}
# marginal effects ATE
qp2 <- glm(realrinc ~ college * (parcol + female + age + I(age^2)), 
           data = d,
           family = "quasipoisson")

avg_slopes(qp2,
           variables = "college",
           type = "link") |> 
  tidy()
```

```{r}
# Manual ATE
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp2, newdata = new_col0) ## predictions for college == 0
p1 <- predict(qp2, newdata = new_col1) ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT Estimate*

```{r}
avg_slopes(qp2,
           variables = "college",
           type = "link",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
# manual ATT
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp2, newdata = new_col0)[d$college == 1] ## predictions for college == 0 for treated
p1 <- predict(qp2, newdata = new_col1)[d$college == 1] ## predictions for college == 1 for treated

# ATT
mean(p1 - p0)
```

### *Manual ATU estimate*

```{r}
# manual ATU
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp2, newdata = new_col0)[d$college == 0] ## predictions for college == 0 for untreated
p1 <- predict(qp2, newdata = new_col1)[d$college == 0] ## predictions for college == 1 for untreated

# ATU
mean(p1 - p0)
```

## 5.3.4 Exercise

```{r}
avg_slopes(qp2,
           variables = "college",
           type = "response") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp2, newdata = new_col0, type = "response") ## predictions for college == 0
p1 <- predict(qp2, newdata = new_col1, type = "response") ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT estimate*

```{r}
# marginal effects ATT/ATU
avg_slopes(qp2,
           variables = "college",
           type = "response",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp2, newdata = new_col0, type = "response")[d$college == 1] 
p1 <- predict(qp2, newdata = new_col1, type = "response")[d$college == 1] 

# ATT estimate
mean(p1 - p0) 
```

### *Manual ATU estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(qp2, newdata = new_col0, type = "response")[d$college == 0] 
p1 <- predict(qp2, newdata = new_col1, type = "response")[d$college == 0] 

# ATU estimate
mean(p1 - p0) 
```

# 5.4 Logistic Regression

```{r}
d <- gss2022 |>
  select(abany, degree, madeg, padeg, sex, age) |> 
  mutate(pardeg = pmax(madeg, padeg, na.rm = TRUE),
         college = if_else(degree >= 3, 1L, 0L),
         parcol = if_else(pardeg >= 3, 1L, 0L),
         female = if_else(sex == 2, 1L, 0L),
         abany = if_else(abany == 1, 1L, 0L)) |>
  select(abany, college, parcol, female, age) |> 
  drop_na()
```

## 5.4.1 Exercise

```{r}
lr1 <- glm(abany ~ college + (parcol + female + age + I(age^2)),
          data = d,
          family = binomial)

# ATE estimate
avg_slopes(lr1,
           variables = "college",
           type = "link") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr1, newdata = new_col0) ## predictions for college == 0
p1 <- predict(lr1, newdata = new_col1) ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT estimate*

```{r}
# marginal effects ATT/ATU
avg_slopes(lr1,
           variables = "college",
           by = "college",
           type = "link") |> 
  tidy()
```

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr1, newdata = new_col0)[d$college == 1] 
p1 <- predict(lr1, newdata = new_col1)[d$college == 1] 

# ATT estimate
mean(p1 - p0) 
```

### *Manual ATU estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr1, newdata = new_col0)[d$college == 0] 
p1 <- predict(lr1, newdata = new_col1)[d$college == 0] 

# ATU estimate
mean(p1 - p0) 
```

## 5.4.2 Exercise

```{r}
# ATE estimate
avg_slopes(lr1,
           variables = "college",
           type = "response") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr1, newdata = new_col0, type = "response") ## predictions for college == 0
p1 <- predict(lr1, newdata = new_col1, type = "response") ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT estimate*

```{r}
# marginal effects ATT/ATU
avg_slopes(lr1,
           variables = "college",
           type = "response",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr1, newdata = new_col0, type = "response")[d$college == 1] 
p1 <- predict(lr1, newdata = new_col1, type = "response")[d$college == 1] 

# ATT estimate
mean(p1 - p0) 
```

### *Manual ATU estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr1, newdata = new_col0, type = "response")[d$college == 0] 
p1 <- predict(lr1, newdata = new_col1, type = "response")[d$college == 0] 

# ATU estimate
mean(p1 - p0) 
```

## 5.4.3 Exercise

```{r}
lr2 <- glm(abany ~ college * (parcol + female + age + I(age^2)),
          data = d,
          family = binomial)

# ATE estimate
avg_slopes(lr2,
           variables = "college",
           type = "link") |> 
  tidy()
```

### *Manual ATE Estimate*

```{r}
# Manual ATE
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr2, newdata = new_col0) ## predictions for college == 0
p1 <- predict(lr2, newdata = new_col1) ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT Estimate*

```{r}
avg_slopes(lr2,
           variables = "college",
           type = "link",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
# manual ATT
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr2, newdata = new_col0)[d$college == 1] 
p1 <- predict(lr2, newdata = new_col1)[d$college == 1] 

# ATT
mean(p1 - p0)
```

### *Manual ATU estimate*

```{r}
# manual ATU
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr2, newdata = new_col0)[d$college == 0] 
p1 <- predict(lr2, newdata = new_col1)[d$college == 0] 

# ATU
mean(p1 - p0)
```

## 5.4.4 Exercise

```{r}
# ATE estimate
avg_slopes(lr2,
           variables = "college",
           type = "response") |> 
  tidy()
```

### *Manual ATE estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr2, newdata = new_col0, type = "response") ## predictions for college == 0
p1 <- predict(lr2, newdata = new_col1, type = "response") ## predictions for college == 1

# ATE estimate
mean(p1 - p0) 
```

### *Manual ATT estimate*

```{r}
# marginal effects ATT/ATU
avg_slopes(lr2,
           variables = "college",
           type = "response",
           by = "college") |> # separately by treatment group
  tidy()
```

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr2, newdata = new_col0, type = "response")[d$college == 1] 
p1 <- predict(lr2, newdata = new_col1, type = "response")[d$college == 1] 

# ATT estimate
mean(p1 - p0) 
```

### *Manual ATU estimate*

```{r}
new_col0 <- d |> mutate(college = 0)
new_col1 <- d |> mutate(college = 1)

p0 <- predict(lr2, newdata = new_col0, type = "response")[d$college == 0] 
p1 <- predict(lr2, newdata = new_col1, type = "response")[d$college == 0] 

# ATU estimate
mean(p1 - p0) 
```
